
---
title: "Data Quality Report"
author: "Team Data Science Process by Microsoft"
date: "`r format(Sys.time(), '%B %d, %Y')`"
output: 
  html_document:
    toc: yes
    toc_float: true
    number_sections: true
    css: style.css
    theme: journal
    highlight: espresso
    fig_width: 7
    fig_height: 6
    fig_caption: true
runtime: shiny

---

# Task Summary


```{r echo = FALSE, message=FALSE, warning=FALSE}

options(warn=-1)

# install required packages
options(repos='http://cran.rstudio.com/')
list.of.packages <- c('Hmisc', 'psych', 'corrgram', 'yaml', 'entropy', 'vcd', 'shiny', 'corrplot', 'scatterplot3d', 'DescTools', 'xtable', 'shinyjs', 'RODBC','parallel','doSNOW','foreach', 'dplyr', 'lubridate')
new.packages <- list.of.packages[!(list.of.packages %in% installed.packages()[,'Package'])]
if(length(new.packages))
  install.packages(new.packages)

# intall knitr version 1.12 
if (!'knitr' %in% installed.packages()[,'Package']){
  knitrurl <- 'http://cran.r-project.org/src/contrib/Archive/knitr/knitr_1.12.tar.gz'
  install.packages(knitrurl, repos=NULL, type='source')
} else if ('1.12' != installed.packages()['knitr','Version']){
  remove.packages('knitr')
  knitrurl <- 'http://cran.r-project.org/src/contrib/Archive/knitr/knitr_1.12.tar.gz'
  install.packages(knitrurl, repos=NULL, type='source')
}

library(yaml)
library(RODBC)
library(foreach)
library(shiny)

# select yaml file
OS_type <- .Platform$OS.type
if (OS_type == 'windows'){
  winDialog(type = 'ok', 'Please select the .yaml file in the next popup window. Click OK to proceed.')
} else{
  print('Please input the path to the .yaml file.')
}
yaml_file <- file.choose()
config <- yaml.load_file(yaml_file)

# data source
if(is.null(config$DataSource) || config$DataSource == 'local'){
  data <- read.csv(config$DataFilePath, header = config$HasHeader, sep =  config$Separator)
} else {
  dbhandle <- odbcDriverConnect(paste0('driver={ODBC Driver 11 for SQL Server};server=',config$Server,';database=',config$Database,';Uid=',config$Username,';Pwd=',config$Password))
  data <- sqlQuery(dbhandle, config$Query)
  odbcClose(dbhandle)
}

# add datetime columns
library(lubridate)

autogen_datetime_columns <- character()
if(!is.null(config$DateTimeColumns)){
  for (dt in names(config$DateTimeColumns)) {
    data[[dt]] <- as.POSIXct(data[[dt]], format = config$DateTimeColumns[[dt]])
    
    new_col_name <- paste0(dt, '_autogen_year')
    data[[new_col_name]] <- year(data[[dt]])
    if (length(unique(na.omit(data[[new_col_name]]))) == 1){
      data[[new_col_name]] <- NULL
    } else{
      autogen_datetime_columns <- c(autogen_datetime_columns, new_col_name)
    }
    

    new_col_name <- paste0(dt, '_autogen_month')
    data[[new_col_name]] <- month(data[[dt]]) 
    if (length(unique(na.omit(data[[new_col_name]]))) == 1){
      data[[new_col_name]] <- NULL
    } else{
      autogen_datetime_columns <- c(autogen_datetime_columns, new_col_name)
    }
    
    new_col_name <- paste0(dt, '_autogen_week')
    data[[new_col_name]] <- week(data[[dt]])
    if (length(unique(na.omit(data[[new_col_name]]))) == 1){
      data[[new_col_name]] <- NULL
    } else{
      autogen_datetime_columns <- c(autogen_datetime_columns, new_col_name)
    }
    
    new_col_name <- paste0(dt, '_autogen_day')
    data[[new_col_name]] <- day(data[[dt]])
    if (length(unique(na.omit(data[[new_col_name]]))) == 1){
      data[[new_col_name]] <- NULL
    } else{
      autogen_datetime_columns <- c(autogen_datetime_columns, new_col_name)
    }
    
    new_col_name <- paste0(dt, '_autogen_wday')
    data[[new_col_name]] <- wday(data[[dt]])
    if (length(unique(na.omit(data[[new_col_name]]))) == 1){
      data[[new_col_name]] <- NULL
    } else{
      autogen_datetime_columns <- c(autogen_datetime_columns, new_col_name)
    }
    
    new_col_name <- paste0(dt, '_autogen_hour')
    data[[new_col_name]] <- hour(data[[dt]])
    if (length(unique(na.omit(data[[new_col_name]]))) == 1){
      data[[new_col_name]] <- NULL
    } else{
      autogen_datetime_columns <- c(autogen_datetime_columns, new_col_name)
    }
    
    new_col_name <- paste0(dt, '_autogen_minute')
    data[[new_col_name]] <- minute(data[[dt]])
    if (length(unique(na.omit(data[[new_col_name]]))) == 1){
      data[[new_col_name]] <- NULL
    } else{
      autogen_datetime_columns <- c(autogen_datetime_columns, new_col_name) 
    }
    
    new_col_name <- paste0(dt, '_autogen_second')
    data[[new_col_name]] <- second(data[[dt]])
    if (length(unique(na.omit(data[[new_col_name]]))) == 1){
      data[[new_col_name]] <- NULL
    } else{
      autogen_datetime_columns <- c(autogen_datetime_columns, new_col_name)
    }
    config$ColumnsToExclude <- c(config$ColumnsToExclude, dt)
  }
}


# Add datetime components to conf$CategoricalColumns
CategoricalColumns <- config$CategoricalColumns
config$CategoricalColumns <- c(config$CategoricalColumns, autogen_datetime_columns)

# detect data types
isNumerical <- sapply(data, is.numeric)
isCategorical <- sapply(data, function(x)length(unique(na.omit(x))) <= nrow(data)/500||length(unique(na.omit(x))) <= 5)
isNumerical <- isNumerical & !isCategorical
colNames <- colnames(data)

# override auto-detected isCategorical with the specified categorical variables in yaml
if(!is.null(config$CategoricalColumns)){
  config$CategoricalColumns <- make.names(config$CategoricalColumns, unique=TRUE)
  for(v in config$CategoricalColumns){
    isCategorical[v] <- TRUE
    isNumerical[v] <- FALSE
  }
}
# override auto-detected isNumerical with the specified numerical variables in yaml
if(!is.null(config$NumericalColumns)){
  config$NumericalColumns <- make.names(config$NumericalColumns, unique = TRUE)
  for(v in config$NumericalColumns){
    isNumerical[v] <- TRUE
    isCategorical[v] <- FALSE
    }
}

# populate config$CategoricalColumns and config$NumericalColumns with detected and specified variables
config$CategoricalColumns <- colNames[isCategorical[colNames] == TRUE]
config$NumericalColumns <- colNames[isNumerical[colNames] == TRUE]

for(v in config$CategoricalColumns)
{
   data[,v] <- as.factor(data[,v])
}  
  
# exclude columns from the report
if(!is.null(config$ColumnsToExclude)){
  config$CategoricalColumns <- config$CategoricalColumns[!config$CategoricalColumns %in% config$ColumnsToExclude]
  config$NumericalColumns <- config$NumericalColumns[!config$NumericalColumns %in% config$ColumnsToExclude]
}

# replace missing values
if(!is.null(config$MissingValueReplaceWith)){
  missingValueReplacement <- config$MissingValueReplaceWith
} else {
  missingValueReplacement <- 0
}

# detect task type
if(is.null(config$Target)){
  taskType <- 'data_exploration'
} else if(isCategorical[config$Target]==FALSE){
  taskType <- 'regression'
} else {
  taskType <- 'classification'
}

# Remove the DateTimeColumns field from the updated YAML file 
config$DateTimeColumns <- NULL

# write updated yaml
if((!is.null(config$DataSource)) && config$DataSource != 'local'){ 
  # if data source is not local file, do not add datetime components into yaml file.
  config$CategoricalColumns <- CategoricalColumns
}
  
config$ColumnsToExclude <- c(config$ColumnsToExclude, names(config$DateTimeColumns))
# output the augmented dataset
if((is.null(config$DataSource) || config$DataSource == 'local') && length(autogen_datetime_columns) > 0){
  library(tools)
  extension <- file_ext(config$DataFilePath)
  path_basename <- file_path_sans_ext(config$DataFilePath)
  augmented_filename <- paste(path_basename, '_dt_components', '.', extension, sep='')
  write.table(data, file = augmented_filename,row.names=FALSE, na="", sep=config$Separator, quote=FALSE)
  config$DataFilePath <- augmented_filename
}
new_yaml_file <- paste0(substr(yaml_file, 1, nchar(yaml_file)-5),'_updated.yaml')
write(as.yaml(config), new_yaml_file)

code = "
#' ---
#' title: 'Data Quality Report'
#' author: 'Team Data Science Process by Microsoft'
#' output: 
#'  html_document:
#'    toc: yes
#' ---
#+ echo=FALSE

options(warn=-1)

# install required packages
options(repos='http://cran.rstudio.com/')
list.of.packages <- c('Hmisc', 'psych', 'corrgram', 'yaml', 'entropy', 'vcd', 'ggvis', 'shiny', 'corrplot', 'scatterplot3d', 'DescTools', 'xtable', 'shinyjs', 'RODBC','parallel','doSNOW','foreach', 'dplyr')
new.packages <- list.of.packages[!(list.of.packages %in% installed.packages()[,'Package'])]
if(length(new.packages))
  install.packages(new.packages)

# intall knitr version 1.12 
if (!'knitr' %in% installed.packages()[,'Package']){
  knitrurl <- 'http://cran.r-project.org/src/contrib/Archive/knitr/knitr_1.12.tar.gz'
  install.packages(knitrurl, repos=NULL, type='source')
} else if ('1.12' != installed.packages()['knitr','Version']){
  remove.packages('knitr')
  knitrurl <- 'http://cran.r-project.org/src/contrib/Archive/knitr/knitr_1.12.tar.gz'
  install.packages(knitrurl, repos=NULL, type='source')
}

library(yaml)
library(RODBC)
library(foreach)

# yaml
yaml_file <- yaml_file_location
config <- yaml.load_file(yaml_file)

# data source
if(is.null(config$DataSource) || config$DataSource == 'local'){
  data <- read.csv(config$DataFilePath, header = config$HasHeader, sep =  config$Separator)
} else {
  dbhandle <- odbcDriverConnect(paste0('driver={ODBC Driver 11 for SQL Server};server=',config$Server,';database=',config$Database,';Uid=',config$Username,';Pwd=',config$Password))
  data <- sqlQuery(dbhandle, config$Query)
  odbcClose(dbhandle)
}

# add datetime columns
library(lubridate)

autogen_datetime_columns <- character()
if(!is.null(config$DateTimeColumns)){
  for (dt in names(config$DateTimeColumns)) {
    data[[dt]] <- as.POSIXct(data[[dt]], format = config$DateTimeColumns[[dt]])
    
    new_col_name <- paste0(dt, '_autogen_year')
    data[[new_col_name]] <- year(data[[dt]])
    if (length(unique(na.omit(data[[new_col_name]]))) == 1){
      data[[new_col_name]] <- NULL
    } else{
      autogen_datetime_columns <- c(autogen_datetime_columns, new_col_name)
    }
    

    new_col_name <- paste0(dt, '_autogen_month')
    data[[new_col_name]] <- month(data[[dt]]) 
    if (length(unique(na.omit(data[[new_col_name]]))) == 1){
      data[[new_col_name]] <- NULL
    } else{
      autogen_datetime_columns <- c(autogen_datetime_columns, new_col_name)
    }
    
    new_col_name <- paste0(dt, '_autogen_week')
    data[[new_col_name]] <- week(data[[dt]])
    if (length(unique(na.omit(data[[new_col_name]]))) == 1){
      data[[new_col_name]] <- NULL
    } else{
      autogen_datetime_columns <- c(autogen_datetime_columns, new_col_name)
    }
    
    new_col_name <- paste0(dt, '_autogen_day')
    data[[new_col_name]] <- day(data[[dt]])
    if (length(unique(na.omit(data[[new_col_name]]))) == 1){
      data[[new_col_name]] <- NULL
    } else{
      autogen_datetime_columns <- c(autogen_datetime_columns, new_col_name)
    }
    
    new_col_name <- paste0(dt, '_autogen_wday')
    data[[new_col_name]] <- wday(data[[dt]])
    if (length(unique(na.omit(data[[new_col_name]]))) == 1){
      data[[new_col_name]] <- NULL
    } else{
      autogen_datetime_columns <- c(autogen_datetime_columns, new_col_name)
    }
    
    new_col_name <- paste0(dt, '_autogen_hour')
    data[[new_col_name]] <- hour(data[[dt]])
    if (length(unique(na.omit(data[[new_col_name]]))) == 1){
      data[[new_col_name]] <- NULL
    } else{
      autogen_datetime_columns <- c(autogen_datetime_columns, new_col_name)
    }
    
    new_col_name <- paste0(dt, '_autogen_minute')
    data[[new_col_name]] <- minute(data[[dt]])
    if (length(unique(na.omit(data[[new_col_name]]))) == 1){
      data[[new_col_name]] <- NULL
    } else{
      autogen_datetime_columns <- c(autogen_datetime_columns, new_col_name) 
    }
    
    new_col_name <- paste0(dt, '_autogen_second')
    data[[new_col_name]] <- second(data[[dt]])
    if (length(unique(na.omit(data[[new_col_name]]))) == 1){
      data[[new_col_name]] <- NULL
    } else{
      autogen_datetime_columns <- c(autogen_datetime_columns, new_col_name)
    }
    config$ColumnsToExclude <- c(config$ColumnsToExclude, dt)
  }
}

# Add datetime components to conf$CategoricalColumns
CategoricalColumns <- config$CategoricalColumns
config$CategoricalColumns <- c(config$CategoricalColumns, autogen_datetime_columns)

# detect data types
isNumerical <- sapply(data, is.numeric)
isCategorical <- sapply(data,function(x)length(unique(na.omit(x)))<=nrow(data)/500||length(unique(na.omit(x)))<=5)
isNumerical <- isNumerical & !isCategorical
colNames <- colnames(data)

# override auto-detected isCategorical with the specified categorical variables in yaml
if(!is.null(config$CategoricalColumns)){
  config$CategoricalColumns <- make.names(config$CategoricalColumns, unique=TRUE)
  for(v in config$CategoricalCoumns){
    isCategorical[v] <- TRUE
    isNumerical[v] <- FALSE
  }
}
# override auto-detected isNumerical with the specified numerical variables in yaml
if(!is.null(config$NumericalColumns)){
  config$NumericalColumns <- make.names(config$NumericalColumns, unique = TRUE)
  for(v in config$NumericalColumns){
    isNumerical[v] <- TRUE
    isCategorical[v] <- FALSE
    }
}

# populate config$CategoricalColumns and config$NumericalColumns with detected and specified variables
config$CategoricalColumns <- colNames[isCategorical[colNames] == TRUE]
config$NumericalColumns <- colNames[isNumerical[colNames] == TRUE]

for(v in config$CategoricalColumns)
{
   data[,v] <- as.factor(data[,v])
} 


# exclude columns from the report
if(!is.null(config$ColumnsToExclude)){
  config$CategoricalColumns <- config$CategoricalColumns[!config$CategoricalColumns %in% config$ColumnsToExclude]
  config$NumericalColumns <- config$NumericalColumns[!config$NumericalColumns %in% config$ColumnsToExclude]
}

# replace missing values
if(!is.null(config$MissingValueReplaceWith)){
  missingValueReplacement <- config$MissingValueReplaceWith
} else {
  missingValueReplacement <- 0
}

# detect task type
if(is.null(config$Target)){
  taskType <- 'data_exploration'
} else if(isCategorical[config$Target]==FALSE){
  taskType <- 'regression'
} else {
  taskType <- 'classification'
}

#' # Task Summary
#+ echo=FALSE
#' - The metadata (location, numerical columns, target, etc.) is - *yaml_file_location*
#' - The data location is - *`r config$DataFilePath`*
#' - The target is - *`r config$Target`*
#' - The task type is - *`r taskType`*.
#' - The numerical variables are - *`r config$NumericalColumns`*
#' - The categorical variables are - *`r config$CategoricalColumns`*
#+ echo=FALSE

"

code <- gsub('yaml_file_location',paste0('"',gsub('\\\\','/',new_yaml_file),'"'), code)
write(code, file = config$RLogFilePath, append = FALSE)

```

- The metadata (categorical columns, numerical columns, target, etc.) is specified in - *`r gsub('\\\\','/',yaml_file)`*
- The data location is - `r config$DataFilePath`.
- The target is - *`r config$Target`*.
- The task type is - *`r taskType`*.
- The numerical variables are - *`r config$NumericalColumns`*.
- The categorical variables are - *`r config$CategoricalColumns`*.

# Data Summary

## Take a peek of the data by showing the top rows of the data
```{r echo = FALSE, message=FALSE, warning=FALSE}

shinyApp(
  
  ui = fluidPage(
  inputPanel(
  numericInput("rows", "Top Rows:", 10)
  ),  
  tableOutput("table"),
  actionButton("submit", label = "Export")
  ),
  
  server = function(input, output) {
    
   r = "
#' # Data Summary
#' ## Take a peek of the data by showing the top rows of the data
#+ echo=FALSE
     head(data, input$rows)
  "
  
    r_code <- reactive({
    gsub("input\\$rows", input$rows, r)
    })
    
    output$table <- renderTable({
      eval(parse(text = r_code()))
    })
    

    observeEvent(input$submit, {
      write(r_code(), file = config$RLogFilePath, append = TRUE)
    })
    
  }
)
```

## The dimensions of the data (Rows, Columns)

```{r echo = FALSE, message = FALSE, warning = FALSE}

 shinyApp(
 
   ui = fluidPage(
   tableOutput("table"),
   actionButton("submit", label = "Export")
   ),
 
   server = function(input, output) {
 
    r = "
#' ## The dimensions of the data (Rows, Columns)
#+ echo=FALSE
dim <- as.data.frame( t(dim(data)))
colnames(dim) <- c('Number of Rows','Number of Columns')
dim
    "
      r_code <- reactive({
        r
    })
      
     output$table <- renderTable({
       eval(parse(text = r_code()))
     })
 
 
     observeEvent(input$submit, {
       write(r_code(), file = config$RLogFilePath, append = TRUE)
     })
   }
 )

```

## Names and types of the columns
```{r echo = FALSE, message=FALSE, warning=FALSE}

 shinyApp(
 
   ui = fluidPage(
   inputPanel(
   numericInput("cols", "Show Columns:", 10)),
   tableOutput("table"),
   actionButton("submit", label = "Export")
   ),
 
   server = function(input, output) {
 
    r = "
#' ## Types of columns
#+ echo=FALSE
# get the data type of each column
 Column_Type <- sapply(data, class)
 Column_Type <- lapply(Column_Type, function(x) paste(x, collapse=' '))
 column_info <- cbind(Column_Name= names(Column_Type), Column_Type)
 rownames(column_info) <- NULL
 column_info[1:min(input$cols,nrow(column_info)),]
    "
      r_code <- reactive({
    gsub("input\\$cols", input$cols, r)
    })
      
     output$table <- renderTable({
       eval(parse(text = r_code()))
     })
 
 
     observeEvent(input$submit, {
       write(r_code(), file = config$RLogFilePath, append = TRUE)
     })
   }
 )

```

## Summarize basic statisics of the data
```{r echo = FALSE, message=FALSE, warning=FALSE}

shinyApp(

  ui = fluidPage(
  tableOutput("table"),  
  actionButton("submit", label = "Export")
  ),

  server = function(input, output) {

   r = "
#' ## Summarize basic statisics of the data
#+ echo=FALSE
summary(data)
   "
    output$table <- renderTable({
      eval(parse(text = r))
    })


    observeEvent(input$submit, {
      write(r, file = config$RLogFilePath, append = TRUE)
    })
  }
)

```

# Dive deeper into each individual variable

## More detailed statistics of each variable

```{r echo = FALSE, message=FALSE, warning=FALSE}
 shinyApp(

   ui = fluidPage(
   inputPanel(
   numericInput("vars", "Show Variables:", 10)),
   verbatimTextOutput("text"),
   actionButton("submit", label = "Export")

   ),

   server = function(input, output) {

    r = "
#' # Dive deeper into each individual variable
#' ## More detailed statistics of each variable
#+ echo=FALSE
 library(Hmisc)
 desc <- Hmisc::describe(as.data.frame(data))
 desc[1:min(input$vars, length(desc))]
    "
      r_code <- reactive({
    gsub("input\\$vars", input$vars, r)
    })

     output$text <- renderPrint({
       eval(parse(text = r_code()))
     })


     observeEvent(input$submit, {
       write(r_code(), file = config$RLogFilePath, append = TRUE)
     })
   }
 )

```


## Visualize the target variable

```{r echo = FALSE, message=FALSE, warning=FALSE}
if(!is.null(config$Target))
{
shinyApp(
  
  ui = fluidPage(
  inputPanel(
  selectInput("target", label = "Target:",  choices = config$Target)
  ),  
  plotOutput("plot"),
  actionButton("submit", label = "Export")
  ),
  
  server = function(input, output) {
    
   r = "
#' # Dive deeper into each individual variable
#' ## Visualize the target variable
#+ echo=FALSE

if(isCategorical[config$Target])
{
    par(mfrow=c(1, 2)) 
    barplot(table(data[[config$Target]]), main = paste('Bar Plot of', config$Target))
    pie(table(data[[config$Target]]), main=paste('Pie Chart of', config$Target))  
}else{
    par(mfrow=c(2,2)) 
    hist(data[[config$Target]], main = paste('Histogram of Target', config$Target), xlab = config$Target)
    # Kernel Density Plot
    d <- density(data[[config$Target]]) # returns the density data 
    plot(d, main = paste('Density Plot of', config$Target)) 
    polygon(d, col='grey', border='blue') 
    qqnorm(data[[config$Target]], main = paste('QQ Plot of Target', config$Target))
    qqline(data[[config$Target]])
    boxplot(data[[config$Target]], main = paste('Boxplot of Target', config$Target))
}
  "
  
    r_code <- reactive({
    gsub("input\\$target", input$target, r)
    })
    
    output$plot <- renderPlot({
      eval(parse(text = r_code()))
    })
    

    observeEvent(input$submit, {
      write(r_code(), file = config$RLogFilePath, append = TRUE)
    })
    
  }
)
} else {
  
      msg = "This is a data exploratory task and it doesn't have a target."
      msg
}
```

## Visualize the numerical variables

You can select the variable from the drop list. 
```{r echo = FALSE, message=FALSE, warning=FALSE}
library(shiny)

shinyApp(
  ui = fluidPage(
    inputPanel(selectInput("numeric", label = "", choices = config$NumericalColumns)),
    plotOutput("plot"), 
    actionButton("submit", label = "Export")
  ),
  
  server = function(input, output) {
    
   r = "
#' ## Visualize the numerical variables
#+ echo=FALSE
  # histogram, density, and QQ plot
  par(mfrow=c(2,2))
  if(length(data[,input$numeric]) >= 5000){
    sampled_data = data[sample(1:nrow(data), 5000, replace=FALSE),]
    normtest <- shapiro.test(sampled_data[[input$numeric]])
  } else{
    normtest <- shapiro.test(data[[input$numeric]])
  }
  
  p.value <- round(normtest$p.value,4)
  if (p.value < 0.05) {
    h0 <- 'rejected.'
    color <- 'red'
  } else {
    h0 <- 'accepted.'
    color <- 'blue'
  }
  hist(data[[input$numeric]], xlab = input$numeric, main = paste('Histogram of', input$numeric))
  
  d <- density(data[[input$numeric]]) 
  plot(d, main = paste('Density Plot of', input$numeric))
  qqnorm(data[[input$numeric]], main = paste('QQ Plot of', input$numeric))
  qqline(data[[input$numeric]])
  boxplot(data[[input$numeric]], main = paste('Boxplot of', input$numeric))
  mtext(paste('Normality test of', input$numeric, h0, '( p-value=', p.value, ')'), side = 3, line = -1, outer = TRUE, fontface = 'italic', col=color, size = 10)
  "
    r_code <- reactive({
	gsub('input\\$numeric', paste0("'",input$numeric,"'"), r)
    })
    
    output$plot <- renderPlot({
      eval(parse(text = r_code()))
    })
    

    observeEvent(input$submit, {
      write(r_code(), file = config$RLogFilePath, append = TRUE)
    })
    
  },
  
  options = list(height = 500)
)
```

## Visualize the categorical variables

You can select the variable from the drop list.

```{r echo = FALSE, message=FALSE, warning=FALSE}
shinyApp(
  
  ui = fluidPage(
    
	  inputPanel(selectInput("categoric", label = "Categorical Variable:",
            choices = config$CategoricalColumns)),
    plotOutput("plot"),
	  actionButton("submit", label = "Export")
  ),
  
  server = function(input, output) {
    
   r = "
#' ## Visualize the categorical variables
#+ echo=FALSE
     # barplot and pie chart
  par(mfrow=c(1,2))
  fhist <- sort(table(data[[input$categoric]]), decreasing = TRUE)
  barplot(table(data[[input$categoric]]), main = paste('Bar Plot of', input$categoric))
  pie(fhist, main=paste('Pie Chart of', input$categoric))
  "
    r_code <- reactive({
	gsub("input\\$categoric", paste0("'",input$categoric,"'"), r)
    })
    
    output$plot <- renderPlot({
      eval(parse(text = r_code()))
    })
    

    observeEvent(input$submit, {
      write(r_code(), file = config$RLogFilePath, append = TRUE)
    })
    
  }
)
```


#  Investigate Multiple Variable Interactions

## Rank variables
Rank variables (numerical and categorical) based on strengths of association with a selected variable. This helps you understand the top dependent variables (grouped by numerical and categorical) of the selected variable.
```{r echo = FALSE, message=FALSE, warning=FALSE}
shinyApp(
  
  ui = fluidPage(
    inputPanel(selectInput("variable", label = "Variable:", choices = if(!is.null(config$Target)) { c(config$Target,config$NumericalColumns[config$NumericalColumns!=config$Target],config$CategoricalColumns[config$CategoricalColumns!=config$Target]) } else {c(config$NumericalColumns, config$CategoricalColumns)}),
  numericInput("top_num", label = "Top Numerical Variables", 5),
  numericInput("top_cat", label = "Top Categorical Variables", 5)),
  actionButton("go", "Go"),
  plotOutput("plot"),
  actionButton("submit", label = "Export")
  ),
  
  server = function(input, output) {
    
   r = "
#' # Investigation on Multiple Variable Interactions
#' ## Rank associated variables
#' This helps you to understand the top dependent variables (grouped by numerical and categorical) of a variable of your choice.
#+ echo=FALSE
    if(nrow(data)>50000) {
      library(dplyr)
      set.seed(9805)
      data <- sample_n(data, min(50000, nrow(data)))
    }

    library(parallel)
    library(doSNOW)
     par(mfrow=c(1,2)) 
  no_cores <- max(detectCores() - 1, 1)
  cluster <- makeCluster(no_cores)
  registerDoSNOW(cluster)
  if(isCategorical[input$variable] == TRUE){
    aov_v <- foreach(i=1:length(config$NumericalColumns),.export=c('config','data'),.packages=c('DescTools'),.combine='c') %dopar%
    {
      get('config')
      get('data')
      if (max(data[[config$NumericalColumns[i]]]) != min(data[[config$NumericalColumns[i]]]))
      {
        col1 <- data[[config$NumericalColumns[i]]]
        index1 <- is.na(col1)
        col1[index1] <- 0
        fit <- aov(col1 ~ data[[input$variable]])
        EtaSq(fit)[1]
        
      } else{
        0
      }
    }
    names(aov_v) <- config$NumericalColumns
    aov_v <- subset(aov_v, names(aov_v)!=input$variable)
    aov_v <- sort(aov_v, decreasing = TRUE)
    barplot(head(aov_v, input$top_num), xlab = 'Eta-squared value', beside=TRUE, main = paste('Top', length(head(aov_v, input$top_num)), 'Associated Numerical Variables'), las=2, cex.axis = 0.7, space=1)
    
    
    
    cramer_v <- foreach(i=1:length(config$CategoricalColumns), .export=c('config','data'), .combine='c') %dopar%
    {
      get('config')
      get('data')
      data[,config$CategoricalColumns[i]] <- factor(data[,config$CategoricalColumns[i]])
      if (nlevels(data[,config$CategoricalColumns[i]]) >= 2) 
      {
        tbl <- table(data[,c(input$variable, config$CategoricalColumns[i])])
        chi2 <- chisq.test(tbl, correct=F)
        sqrt(chi2$statistic / sum(tbl))
      } else{
        0
      }
      
    }
    names(cramer_v) <- config$CategoricalColumns
    cramer_v <- subset(cramer_v, names(cramer_v)!=input$variable)
    cramer_v <- sort(cramer_v, decreasing = TRUE)
    barplot(head(cramer_v, input$top_cat), xlab = 'Cramer\\'s V', beside=TRUE, main = paste('Top', length(head(cramer_v, input$top_cat)), 'Associated Categorical Variables'), las=2, cex.axis = 0.7)
    
    
  } else{
    if(length(config$NumericalColumns)>=2){
      cor <- cor(data[,input$variable], data[,config$NumericalColumns], method = 'pearson')
      cor=cor[1,]
      names(cor) <- config$NumericalColumns
      cor <- subset( cor, names(cor) != input$variable)
      cor_s <- cor*cor
      names(cor_s) <- names(cor)
      cor_s <- sort(cor_s, decreasing = TRUE)
      barplot(head(cor_s, input$top_num), xlab = 'R-squared (squared correlation)', beside=TRUE, main = paste('Top', length(head(cor_s, input$top_num)), 'Associated Numerical Variables'), las=2, cex.axis = 0.7)
    }
    
    aov_v <- foreach(i=1:length(config$CategoricalColumns), .export=c('config', 'data'), .packages=c('DescTools'), .combine='c') %dopar%
    {
      get('config')
      get('data')
      catCols <- config$CategoricalColumns
      numCols <- config$NumericalColumns
      x <- factor(data[[catCols[i]]])
      if (nlevels(x) >= 2 & nlevels(x) <=500)
      {
        fit <- aov(data[[input$variable]]~ x)
        EtaSq(fit)[1]
      } else{
        0
      }
      
    }
    names(aov_v) <- config$CategoricalColumns
    aov_v <- subset(aov_v, names(aov_v)!=input$variable)
    aov_v <- sort(aov_v, decreasing = TRUE)
    barplot(head(aov_v, input$top_cat), xlab = 'Eta-squared value', beside=TRUE, main = paste('Top', length(head(aov_v, input$top_cat)), 'Associated Categorical Variables'), las=2, cex.axis = 0.7)
  }
  stopCluster(cluster)
  "

    r_code <- eventReactive(input$go, {
      r = gsub("input\\$variable", paste0("'",input$variable,"'"), r)
  	  r = gsub("input\\$top_num", input$top_num, r)
	    gsub("input\\$top_cat", input$top_cat, r)
	    })
    output$plot <- renderPlot({
      eval(parse(text = r_code()))
    })
    

    observeEvent(input$submit, {
      write(r_code(), file = config$RLogFilePath, append = TRUE)
    })
    
  }
)

```

## Visualize interactions between two categorical variables

You can select the two variables from the drop lists.
```{r echo = FALSE, message=FALSE, warning=FALSE}
shinyApp(
  
  ui = fluidPage(
  inputPanel(
  selectInput("categoric1", label = "Categorical Variable:", choices = if (!is.null(config$Target) && isCategorical[config$Target]) c(config$Target, config$CategoricalColumns[config$CategoricalColumns!=config$Target]) else config$CategoricalColumns),
  selectInput("categoric2", label = "Categorical Variable:", choices = if (!is.null(config$Target) && isCategorical[config$Target]) c(config$Target, config$CategoricalColumns[config$CategoricalColumns!=config$Target]) else config$CategoricalColumns)),
  plotOutput("plot"),
  actionButton("submit", label = "Export")
  ),
  
  server = function(input, output) {
    
   r = "
#' ## Visualize interactions between two categorical variables
#+ echo=FALSE
  library(vcd)
  par(mfrow=c(1,1)) 
  mosaicplot(table(data[[input$categoric1]], data[[input$categoric2]]),  shade=TRUE, xlab=input$categoric1, ylab=input$categoric2, main=paste(input$categoric1,'VS', input$categoric2))
   "
  
    r_code <- reactive({
      r = gsub("input\\$categoric1", paste0("'",input$categoric1,"'"), r)
      gsub("input\\$categoric2", paste0("'",input$categoric2,"'"), r)
    })
    
    output$plot <- renderPlot({
      eval(parse(text = r_code()))
    })
    

    observeEvent(input$submit, {
      write(r_code(), file = config$RLogFilePath, append = TRUE)
    })
    
  }
)
```

## Visualize interactions between two numerical variables

You can select the two variables from the drop lists.
```{r echo = FALSE, message=FALSE, warning=FALSE}
shinyApp(
  
  ui = fluidPage(
  inputPanel(
  selectInput("numeric1", label = "Numerical Variable:", choices = if (!is.null(config$Target) && isNumerical[config$Target]) c(config$Target, config$NumericalColumns[config$NumericalColumns!=config$Target]) else config$NumericalColumns),
  selectInput("numeric2", label = "Numerical Variable:", choices = if (!is.null(config$Target) && isNumerical[config$Target]) c(config$Target, config$NumericalColumns[config$NumericalColumns!=config$Target]) else config$NumericalColumns)),  
  plotOutput("plot"),
  actionButton("submit", label = "Export")
  ),
  
  server = function(input, output) {
    
   r = "
#' ## Visualize interactions between two numerical variables
#+ echo=FALSE
  par(mfrow=c(1,1)) 
  correlation <- round(cor(data[[input$numeric1]], data[[input$numeric2]]),4)
  plot(data[[input$numeric1]], data[[input$numeric2]], xlab = input$numeric1, ylab=input$numeric2, main = paste(input$numeric1, 'VS', input$numeric2, 'correlation =', correlation))
  # regression line (y~x)
  abline(lm(as.formula(paste(input$numeric1,'~',input$numeric2)), data =data), col='red')  
  # lowess line (x,y) 
  lines(lowess(data[[input$numeric1]], data[[input$numeric2]]), col='blue') 
   "
  
    r_code <- reactive({
      r = gsub("input\\$numeric1", paste0("'",input$numeric1,"'"), r)
	    gsub("input\\$numeric2", paste0("'",input$numeric2,"'"), r)
    })
    
    output$plot <- renderPlot({
      eval(parse(text = r_code()))
    })
    

    observeEvent(input$submit, {
      write(r_code(), file = config$RLogFilePath, append = TRUE)
    })
    
  }
)
```

## Calculate the correlations between numerical variables

```{r echo = FALSE, message=FALSE, warning=FALSE}

if(length(config$NumericalColumns)>=2)
{
shinyApp(
  
  ui = fluidPage(
    inputPanel(
      selectInput("cormethod", label = "Correlation Method:", choices = c("pearson","kendall", "spearman")),
      selectInput("cororder", label = "Order:", choices = c("AOE","FPC", "hclust", "alphabet")),  
      selectInput("corvismethod", label = "Shape:", choices = c("circle","square","ellipse","number","shade","color","pie")),
      selectInput("corvistype", label = "Layout:", choices = c("full","upper", "lower"))),  
    plotOutput("plot"),
    actionButton("submit", label = "Export")
  ),
  
  server = function(input, output) {
    
    r = "
#' ## Calculate the correlations (pearson, kendall, or spearman) between numerical variables
#+ echo=FALSE
    library(corrgram)
    library(corrplot)
    par(mfrow=c(1,1))
    data[is.na(data)] <- missingValueReplacement
    c <- cor(data[,config$NumericalColumns], method = input$cormethod)
    c[is.na(c)] <- missingValueReplacement
    corrplot(c, method=input$corvismethod, order = input$cororder, insig = 'p-value',  sig.level=-1, type = input$corvistype)
    "
    
    r_code <- reactive({
      r = gsub("input\\$cormethod", paste0("'",input$cormethod,"'"), r)
      r = gsub("input\\$cororder", paste0("'",input$cororder,"'"), r)
      r = gsub("input\\$corvismethod", paste0("'",input$corvismethod,"'"), r)
      gsub("input\\$corvistype", paste0("'",input$corvistype,"'"), r)
    })
    output$plot <- renderPlot({
      eval(parse(text = r_code()))
    })
    
    observeEvent(input$submit, {
      write(r_code(), file = config$RLogFilePath, append = TRUE)
    })
  }
)
}

```

## Visualize interactions between numerical and categorical variables via box plots

X axis is the level of categorical variables. This helps you to understand whether the distribution of the numeric variable is significantly different at different levels of the categorical variable. 

We test hypothesis 0 (h0) that the numeric variable has the same mean values across the different levels of the categorical variable. 
```{r echo = FALSE, message=FALSE, warning=FALSE}
library(vcd)

shinyApp(
  
  ui = fluidPage(
    inputPanel(
      selectInput("numeric3", label = "Numeric Variable:", choices = if (!is.null(config$Target) && isNumerical[config$Target]) c(config$Target, config$NumericalColumns[config$NumericalColumns!=config$Target]) else config$NumericalColumns),
      selectInput("categoric3", label = "Categorical Variable:", choices = if (!is.null(config$Target) && isCategorical[config$Target]) c(config$Target,config$CategoricalColumns[config$CategoricalColumns!=config$Target]) else config$CategoricalColumns)
    ),   
    plotOutput("plot"),
    actionButton("submit", label = "Export")
  ),
  
  server = function(input, output) {
    
    r = "
#' ## Visualize interactions between numeric and categorical variables via box plots
#' X axis is the level of categorical variables. This helps you to understand whether the distribution of the numeric variable is significantly different at different levels #' of the categorical variable. 
#' We test hypothesis 0 (h0) that the numeric variable has the same mean values across the different levels of the categorical variable. 
#+ echo=FALSE
    
    par(mfrow=c(1,1)) 
    fit <- aov(data[[input$numeric3]] ~ data[[input$categoric3]])
    test_results <- drop1(fit,~.,test='F')
    p_value <- round(test_results[[6]][2],4)
    if (p_value < 0.05){
    h0 <- 'Rejected'
    color <- 'red'
    } else{
    h0 <- 'Accepted'
    color <- 'blue'
    }
    f <- as.formula(paste(input$numeric3,'~',input$categoric3))
    boxplot(f, data= data, xlab = input$categoric3, ylab=input$numeric3)
    title(main=paste('h0', h0, '( p-value=', p_value, ')'), col.main=color)
    "
    
    r_code <- reactive({
      r = gsub("input\\$numeric3", paste0("'",input$numeric3,"'"), r)
      gsub("input\\$categoric3", paste0("'",input$categoric3,"'"), r)
    })
    
    output$plot <- renderPlot({
      eval(parse(text = r_code()))
    })
    
    
    observeEvent(input$submit, {
      write(r_code(), file = config$RLogFilePath, append = TRUE)
    })
    
  }
)
```

## Project numerical variables to principal components, and visualize

```{r echo = FALSE, message=FALSE, warning=FALSE}

if(length(config$NumericalColumns)>=2)
{
library(scatterplot3d)
data[is.na(data)] <- missingValueReplacement
x <- apply(data[,config$NumericalColumns],2,min)
y <- apply(data[,config$NumericalColumns],2,max)
index <- x == y
nonConstantNames <- config$NumericalColumns[!index]

x <- data[,nonConstantNames]
sigma <- cor(x)
sigma_eigen <- eigen(sigma)
sigma_values <- sigma_eigen$values
index <- sigma_values < 0 
if (sum(index) > 0)
{
  sigma_values[index] <- 0
}
sum_variance <- sum(sigma_values^2)
x <- scale(x)
loadings <- x %*% sigma_eigen$vectors
p.variance.explained <- sigma_values^2/sum_variance
p.variance.cumsum <- cumsum(p.variance.explained)*100

x <- loadings[,1]
y <- loadings[,2]
z <- loadings[,3]
num_numericvars <- length(nonConstantNames)

r = "
#+ echo=FALSE

if(nrow(data)>50000) {
    library(dplyr)
    set.seed(9805)
    data <- sample_n(data, min(50000, nrow(data)))
}

library(scatterplot3d)
data[is.na(data)] <- missingValueReplacement
x <- apply(data[,config$NumericalColumns],2,min)
y <- apply(data[,config$NumericalColumns],2,max)
index <- x == y
nonConstantNames <- config$NumericalColumns[!index]

x <- data[,nonConstantNames]
sigma <- cor(x)
sigma_eigen <- eigen(sigma)
sigma_values <- sigma_eigen$values
index <- sigma_values < 0 
if (sum(index) > 0)
{
  sigma_values[index] <- 0
}
sum_variance <- sum(sigma_values^2)
x <- scale(x)
loadings <- x %*% sigma_eigen$vectors
p.variance.explained <- sigma_values^2/sum_variance
p.variance.cumsum <- cumsum(p.variance.explained)*100

num_numericvars <- length(nonConstantNames)
"
write(r, file = config$RLogFilePath, append = TRUE)

shinyApp(
  
  ui = fluidPage(
    inputPanel(
      selectInput("catvar", label = "Color by categorical variable", choices = if(!is.null(config$Target) && isCategorical[config$Target]) c(config$Target,config$CategoricalColumns[config$CategoricalColumns!=config$Target]) else config$CategoricalColumns),
      selectInput("xvar", label = "Principal component at x axis", choices = c(1:num_numericvars)),
      selectInput("yvar", label = "Principal component at y axis", choices = c(2,1,3:num_numericvars))
    ),  
    plotOutput("plot"),
    actionButton("submit", label = "Export")
  ),
  
  server = function(input, output) {
    
    r = "
#' ## Project numeric variables to principal components, and visualize
#+ echo=FALSE
par(mfrow=c(1,2))
# plot percentage of variance explained for each principal component
ylimit <- ceil(max(p.variance.explained)*100/5)*5
barplot(100*p.variance.explained, las=2, xlab='Principal Components', ylab='% Variance Explained', xaxt='n', yaxt='n')
axis(2, pretty(c(0,ylimit)), col='blue')
box()
par(new=TRUE)
plot(1:num_numericvars, p.variance.cumsum, type='l', col='black', ylab='', xlab='', las=1, axes=FALSE, ylim=c(0,100), xaxt='n')
axis(4, pretty(c(0,100)), col='black',col.axis='black',las=1, axes=F)
num_pcs_80 <- sum(p.variance.cumsum <= 80)
num_pcs_90 <- sum(p.variance.cumsum <= 90)
num_pcs_95 <- sum(p.variance.cumsum <= 95)
text(num_numericvars/10*3, 80, paste('80% by', num_pcs_80, 'pcs'))
text(num_numericvars/10*3, 85, paste('90% by', num_pcs_90, 'pcs'))
text(num_numericvars/10*3, 90, paste('95% by', num_pcs_95, 'pcs'))
data[[input$catvar]] <- factor(data[[input$catvar]])
plot(loadings[,as.numeric(input$xvar)], loadings[,as.numeric(input$yvar)], type='p', pch=20, col=as.numeric(data[[input$catvar]]), xlab=paste('PC', input$xvar), ylab=paste('PC', input$yvar, sep=''))
legend('topright', cex=.8,  legend = levels(data[[input$catvar]]), fill = 1:nlevels(data[[input$catvar]]), merge = F, bty = 'n')
    "
    
    r_code <- reactive({
      r = gsub("input\\$catvar", paste0("'",input$catvar,"'"), r)
      r = gsub("input\\$xvar", paste0("'",input$xvar,"'"), r)
      gsub("input\\$yvar", paste0("'",input$yvar,"'"), r)
    })
    
    output$plot <- renderPlot({
      eval(parse(text = r_code()))
    })
    
    
    observeEvent(input$submit, {
      write(r_code(), file = config$RLogFilePath, append = TRUE)
    })
    
  }
)

}
```

```{r echo = FALSE, message=FALSE, warning=FALSE}
if(length(config$NumericalColumns)>=3)
{
shinyApp(
  
  ui = fluidPage(
    inputPanel(
      selectInput("cat1", label = "Color by categorical variable", choices = if (!is.null(config$Target) &&isCategorical[config$Target]) c(config$Target, config$CategoricalColumns[config$CategoricalColumns!=config$Target]) else config$CategoricalColumns),
      selectInput('pc1', label = 'PC at x axis', choices = c(1:num_numericvars), selected = '1'),
      selectInput('pc2', label = 'PC at y axis', choices = c(1:num_numericvars), selected = '2'),
      selectInput('pc3', label = 'PC at z axis', choices = c(1:num_numericvars), selected = '3'),
      sliderInput("angle", label = "View Angle", min = -180, max = 180, value = 40)
    ),  
    plotOutput("plot"),
    actionButton("submit", label = "Export")
  ),
  
  server = function(input, output) {
    
    r = "
#' ## Project numeric variables to principal components, and visualize
#+ echo=FALSE
    x <- loadings[,as.numeric(input$pc1)]
    y <- loadings[,as.numeric(input$pc2)]
    z <- loadings[,as.numeric(input$pc3)]
    data[[input$cat1]] <- factor(data[[input$cat1]])
    par(mfrow=c(1,1))
    DF <- data.frame(x = x, y = y, z = z, group = data[[input$cat1]])
    # create the plot, you can be more adventurous with colour if you wish
    s3d <- with(DF, scatterplot3d(x, y, z, xlab=paste0('PC',input$pc1), ylab=paste0('PC',input$pc2), zlab=paste0('PC', input$pc3), color = as.numeric(group), pch = 19, angle = as.numeric(input$angle)))
    legend('topleft', cex=.8,  legend = levels(data[[input$cat1]]), fill = 1:nlevels(data[[input$cat1]]), merge = F, bty = 'n')
    "
    
    r_code <- reactive({
      r = gsub("input\\$cat1", paste0("'",input$cat1,"'"), r)
      r = gsub("input\\$angle", paste0("'",input$angle,"'"), r)
      r = gsub("input\\$pc1", paste0("'",input$pc1,"'"), r)
      r = gsub("input\\$pc2", paste0("'",input$pc2,"'"), r)
      gsub("input\\$pc3", paste0("'",input$pc3,"'"), r)
    })
    
    output$plot <- renderPlot({
      eval(parse(text = r_code()))
    })
    
    observeEvent(input$submit, {
      write(r_code(), file = config$RLogFilePath, append = TRUE)
    })
    
  }
)
}

```

# Final Report

```{r, echo = FALSE, message=FALSE}
library(shinyjs)
shinyApp(
  
  ui = fluidPage(
    shinyjs::useShinyjs(),
    inputPanel(
      actionButton('generateReport', 'Generate Report'),
      shinyjs::hidden(actionButton('viewReport', 'View Report'))
    ) 
  ),
  
  server = function(input, output) {
    
    observeEvent(input$generateReport, {
      shinyjs::disable('viewReport')
      withProgress(message = 'Report Generation in progress', value = 0, {
        rmarkdown::render(config$RLogFilePath, clean = FALSE)
      })
      shinyjs::show('viewReport')
      shinyjs::enable('viewReport')
    })
    observeEvent(input$viewReport, {
      browseURL(file.path(getwd(), paste0(substr(config$RLogFilePath, 1, nchar(config$RLogFilePath)-1),'html')))
    })
    
  }
)
```
